---
title: "QMBR Seminar 1"
author: "Instructor: E.D. Starshov (e.starshov@gsom.spbu.ru)"
topics: "Topic 1. Data gathering and analysis; Topic 2. Hypothesis testing"
data: "BEEPS_12_inclass.sav"
---

# Quantitative methods of business research. Seminar 1. #

The goal of the first seminar is to get familiar with RStudio interface and to learn how to process the data there. You will familiarize yourself with R functions which allow you to check if model assumptions hold and test different hypothesis. 
We will use a pre-made dataset. It is real and interesting data about business environment and enterprise performance in East European countries that was originally collected by the European Bank for Reconstruction and Development (EBRD) and the World Bank Group (the World Bank). 

For the seminar we will follow the script below, and it will be similar for the home task (although the datafile and specific questions will be slightly different).



## Q0. Installing and loading required packages and data for the analysis

```{r message=FALSE, warning=FALSE}
# Install packages
install.packages("haven")
install.packages("psych")
install.packages("car")
install.packages("DescTools")
```


```{r message=FALSE, warning=FALSE}
# Activate packages
library(haven) # Loading SPSS files
library(psych) # describe
library(car) # qqPlot
library(DescTools) # DunnettTest
```


```{r}
# Set working directory
setwd("C:/Users/user/OneDrive - St.Petersburg State University/Документы")

# Load the data
beeps12 <- read_sav("D:/GSOM/GraphicalDataAnalysis/Quantitave Methods/data/beeps12.sav") # Specify your own path
View(beeps12)
```



## Q1. Getting acquainted with the data.

```{r}
# b)	Companies from which countries are selected?	
table(beeps12$cname) # categories and their size
```


```{r}
# c)	When was this data collected?	
table(beeps12$year)
```


```{r}
# d)	What does variable own tell you about? What is its type? What is the measurement level?
table(beeps12$own) 
```


```{r}
str(beeps12$own) # variable structure
```


```{r}
attributes(beeps12$own) # variable attributes
```



## Q2. Analyzing for outliers and building graphs


```{r}
# Histogram
hist(beeps12$employ, main = "Histogram of number of employees")
```



```{r}
# Build a simple boxplot
boxplot(beeps12$employ)
```


```{r}
# Build a horizontal boxplot
boxplot(beeps12$employ, horizontal = TRUE, main = "Box plot of number of employess")
```

Three ways to save graphs:
1. Print screen
2. Copypaste the code to the console and run it => Export => Save as image
3. Using the following code


```{r}
# Code for saving the plot to a png file
setwd("C:/Users/user/OneDrive - St.Petersburg State University/Документы")

png(file = "beeps12employBoxplot.png", width = 600, height = 350)
boxplot(beeps12$employ, horizontal = TRUE, main = "Box plot of number of employees")
dev.off()
```


```{r}
# Look at the outlier values
boxplot(beeps12$employ, plot = FALSE)$out
```



```{r}
# Mean of a single variable
mean(beeps12$employ)
```


```{r}
# Standard deviation of a single variable
sd(beeps12$employ)
```


```{r}
# Use "describe" function from "psych" package to get both
describe(beeps12$employ)
```


### Removing outliers


```{r}
# Save outlier values as a separate object
outliers <- boxplot(beeps12$employ, plot = FALSE)$out
```


```{r}
# Create a copy of the dataset without outliers
beeps12clean <- beeps12[-which(beeps12$employ %in% outliers), ]
```


For beginners it can be easier to find the smallest value among outliers and filter the dataset.

```{r}
# Find the smallest outlier
min(boxplot(beeps12$employ, plot = FALSE)$out)
```


```{r}
# Save filtered dataset 
beeps12clean <- beeps12[beeps12$employ < 76, ]
```


```{r}
# Build a simple boxplot
boxplot(beeps12clean$employ,  main = "Box plot of number of employees")
```


### QUESTION TO DISCUSS: What else should I consider when choosing the threshold value?

Let us consider 125 as the threshold value.

```{r}
# Save filtered dataset 
beeps12clean <- beeps12[beeps12$employ <= 125, ]
```


```{r}
# Build a simple boxplot
boxplot(beeps12clean$employ)
```


```{r}
# Look at new descriptive statistics
describe(beeps12clean$employ)
```



## Q3. Normality check

### QUESTION TO DISCUSS: What is central limit theorem? If we have more than 30 observations, does it mean that our distribution is normal?

Let’s now test the assumption for future tests – the normality assumption. 
Let’s first do that using graph. 

```{r}
# Q-Q plot
qqnorm(beeps12clean$employ, main = "QQ plot of number of employees")
qqline(beeps12clean$employ, col = "steelblue", lwd = 2)
```


```{r}
# A simpler code using a command from "car" package
qqPlot(beeps12clean$employ, main = "QQ plot of number of employees")
```


Evaluate Skewness and Kurtosis for the variable employ. 


```{r}
# You can evalueate skewness and kurtosis in "describe" output
describe(beeps12clean$employ)
```


*More info on testing normality on skewness and kurtosis: https://imaging.mrc-cbu.cam.ac.uk/statswiki/FAQ/Simon*

Another way to test normality is to use Kolmogorov Smirnov test. Do this again for the variable employ. 

```{r}
# Kolmogorov-Smirnov test
ks.test(beeps12clean$employ, "pnorm")
```

```{r}
# Shapiro-Wilk test
shapiro.test(beeps12clean$employ)
```
It might be worth to reconstruct the histogram.

```{r}
# Histogram
hist(beeps12clean$employ, main = "Histogram of number of employees")
```


Let’s now separately test hypotheses about different countries. To do that you need to split the file into groups by country name.

```{r}
# Create two separate R objects corresponding to each subsample
beeps12clean1 <- beeps12clean[beeps12clean$cname == "Country 1", ]
beeps12clean2 <- beeps12clean[beeps12clean$cname == "Country 2", ]
```



## Q4. Use an appropriate test and check if the mean number of full-time employees working in Country 1 in 2012 is equal to 25 people. How about Country 2?

### First describe variable employ for two different countries (check for mean, median, standard deviation). 

```{r}
# Describe two subsamples

## Country 1
describe(beeps12clean1$employ)
```


```{r}
## Country 2
describe(beeps12clean2$employ)
```


```{r}
# Build box plots for both countries
boxplot(employ ~ cname, data = beeps12clean)
```

```{r}
# Build histograms for both countries

## Country 1
hist(beeps12clean1$employ)
```


```{r}
## Country 2
hist(beeps12clean2$employ)
```

```{r}
# Kolmogorov-Smirnov test

## Country 1
ks.test(beeps12clean1$employ, "pnorm")
```


```{r}
## Country 2
ks.test(beeps12clean2$employ, "pnorm")
```


```{r}
# Levene's test
leveneTest(employ ~ cname, data = beeps12clean)
```

NOTE: If there is a stat. difference, categorical variable can be later used as control variable, e.g., in regression analysis.


### Now let's move on to the tests.

Normality assomption can be relaxed for one-sample t-tests (when we compare the mean to a specific value).

```{r}
# (Parametric) t-test for country 1
t.test(beeps12clean1$employ, mu = 25)
```


```{r}
# (Non-parametric) Mann-Whitney test for country 2
t.test(beeps12clean2$employ, mu = 25)
```
*More info on t-test and its types: https://www.scribbr.com/statistics/t-test/*


## Q5. Use an appropriate test and test if the average number of full-time employees working in Country 1 in 2012 and the similar number in Country 2 are equal.


```{r}
# (Parametric) T-test for comparing means of both countries
t.test(beeps12clean1$employ,  beeps12clean2$employ, var.equal = TRUE)
```


```{r}
# Mann-Whitney non-parametric test for comparing means of both countries
wilcox.test(beeps12clean1$employ, beeps12clean2$employ)
```


```{r}
# It might be worth to reconstruct the box plots for both countries
boxplot(employ ~ cname, data = beeps12clean)
```

## Q6. Use an appropriate test and test if the average number of full-time employees working in 2012 doesn't differ for companies with different ownership. 


### First test ANOVA assumptions

#### Variance homogeneity

```{r}
# Levene's test
leveneTest(employ ~ own, data = beeps12clean)
```

This happens because R treats own variable as quantitative. We need to convert it into categorical:

```{r}
# Factorize variable
beeps12clean$own <- as.factor(beeps12clean$own)
```


```{r}
# Levene's test
leveneTest(employ ~ own, data = beeps12clean)
```


#### Normal distribution

```{r}
# Kolmogorov - Smirnov test and histograms

## 1
ks.test(beeps12clean$employ[beeps12clean$own == 1], "pnorm") # Instead of creating new objects we may use filters
hist(beeps12clean$employ[beeps12clean$own == 1])
```


```{r}
## 2
ks.test(beeps12clean$employ[beeps12clean$own == 2], "pnorm")
hist(beeps12clean$employ[beeps12clean$own == 2])
```


```{r}
## 3
ks.test(beeps12clean$employ[beeps12clean$own == 3], "pnorm")
hist(beeps12clean$employ[beeps12clean$own == 3])
```


Which test should we use?
* Equal var & normality => ANOVA
* Non-equal var => Welch ANOVA
* Non-normality => Kruskal-Wallis H Test

Let's do all and check the results.

```{r}
# ANOVA: statistical test on equality of the population means 

# Compute the analysis of variance
res.aov <- aov(employ ~ own, data = beeps12clean)

# Summary of the analysis
summary(res.aov)
```


```{r}
# Welch's ANOVA
oneway.test(employ ~ own, data = beeps12clean, var.equal = TRUE)
```


```{r}
# Kruskal-Wallis test
kruskal.test(employ ~ own, data = beeps12clean)
```

### Post-hoc tests

Which test should we use?
* Equal var: Tuckey test
* Non-equal var: Dunnett's test


```{r}
# Tuckey test
TukeyHSD(res.aov)
```


```{r}
# Dunnett's test
DunnettTest(x = beeps12clean$employ, g = beeps12clean$own)
```


## Q7. Creating new variables

Let’s modify some variables to make them more useful for analysis. 

```{r}
# Create a new variable by adding up two
beeps12$percexport <- beeps12$q14b + beeps12$q14c 
```

Next we turn to another Transform procedure subroutine, called Recode. This procedure allows us to create a new variable by recoding scores on another variable. 

Use this procedure to convert the continuous variable percexport into the categorical variable exporter. 
•	If a company’s percexport score is 50 or lower, then that score will be transformed to a value of 1 on the new exporter variable (i.e., company will be placed in the low exporting category)
•	If a company’s percexport score is 51 or higher, then that score will be transformed a value of 2 on the new exporter variable (i.e., company will be placed in the low exporting category)


```{r}
# Create a new categorical variable
beeps12$exporter <- as.factor(ifelse(beeps12$percexport <= 50, 1, 2))
```


```{r}
# Chack the proportions of categories
table(beeps12$exporter)
```

Is everything all right with these numbers?


```{r}
# Look at the percexport variable
table(beeps12$percexport)
```


We have 9 observations with negative value which is nonsense!
But more importantly, we have non-importers, which should be categorized separately.
Let's delete outliers and recode the exporter variable:


```{r}
# Recode exporter variable
beeps12new <- beeps12[beeps12$percexport >= 0, ]
beeps12new$exporter <- as.factor(ifelse(beeps12new$percexport == 0, 0, 
                                        ifelse(beeps12new$percexport <= 50, 1, 2)))
```


```{r}
# Check new categories and their size
table(beeps12new$exporter)
```


## Q8. Checking on total sales (independet work)

Now let’s check total sales variable: is it normally distributed, are there any outliers, think of some tests that can be interesting for sales comparison in this dataset (2 different tests).  


```{r}
### Type your code here
```

